Verification of the CVM algorithm with a New Recursive Analysis Technique

### 2025-02-10 paper/sec/original_proof.tex
Paper: Added related work section. (WIP)

In \cref{alg:cvm_simul} we describe the new algorithm, that behaves exactly as the original one (\cref{alg:cvm}), but performs indepenent coin flips.
Note that the function $\mathrm{last{\textunderscore}index}$ returns the index of the last-occurence of an element in the sequence, before the current loop iteration.
It should be noted that the algorithm keeps track of the number of subsampling iterations $k$, instead of the value $p = 2^{-k}$ as the original algorithm does.
%Like, we did in \cref{sec:invariants}, we ignore the second check, whether the subsampling operation succeeded.
%As we explained there, the total variational difference between these two variants is $\frac{\delta}{2}$.
%
\begin{algorithm}[h!]
	\caption{Modified CVM algorithm with independent coin flips.}\label{alg:cvm_simul}
	\begin{algorithmic}[1]
  \Require Stream elements $a_1,\dots,a_l$, $0 < \varepsilon$, $0 < \delta < 1$.
  \State $\chi \gets \{\}, k \gets 0, n = \ceil*{\frac{12}{\varepsilon^2} \ln{(\frac{6l}{\delta})} }$
  \State $b[i,j] \getsr \Ber(1/2)$ for $i,j \in \{1,\cdots,l\}$ \Comment perform $l^2$ unbiased independent coin flips
  \For{$i \gets 1$ to $l$}
    \If{$b[i,1]=b[i,2]=\cdots=b[i,k]=1$}
      \State $\chi \gets \chi \cup \{a_i\}$
    \Else
      \State $\chi \gets \chi - \{a_i\}$
    \EndIf
    \If{$|\chi| = n$}
      \State $\chi \gets \{a \in \chi \;|\; b[\mathrm{last{\textunderscore}index}(a),k+1] = 1\}$
      \State $k \gets k+1$
    \EndIf
    \If{$|\chi| = n$}
      \State \Return $\bot$
    \EndIf
  \EndFor
  \State \Return $2^k |\chi|$ \Comment estimate cardinality of $A$
  \end{algorithmic}
\end{algorithm}
To roughly explain, how tail bounds can be derived for this new variant:
It is possible to union bound the probability that the estimate exceeds the desired interval, with the probability of the event in conjunction with a specific value of $k$, which can be bounded by the probability that the number of stream elements with whose associated coin flips start with $k$ ones, is outside of $2^{-k} |A| (1 \pm \varepsilon)$.
This is explained in more detail by Chakraborty et al.~\cite{chakraborty2022}.

>>>>

In \cref{alg:cvm_simul} we describe the new algorithm, that behaves exactly as the original one (\cref{alg:cvm}), but performs indepenent coin flips.
Note that the function last{\textunderscore}index returns the index of the last-occurence of an element in the sequence, before the current loop iteration.
It should be noted that the algorithm keeps track of the number of subsampling iterations $k$, instead of the value $p = 2^{-k}$ as the original algorithm does.
%Like, we did in \cref{sec:invariants}, we ignore the second check, whether the subsampling operation succeeded.
%As we explained there, the total variational difference between these two variants is $\frac{\delta}{2}$.
%
\begin{algorithm}[h!]
	\caption{Modified CVM algorithm with independent coin flips.}\label{alg:cvm_simul}
	\begin{algorithmic}[1]
  \Require Stream elements $a_1,\dots,a_l$, $0 < \varepsilon$, $0 < \delta < 1$.
  \State $\chi \gets \{\}, k \gets 0, n = \ceil*{\frac{12}{\varepsilon^2} \ln{(\frac{6l}{\delta})} }$
  \State $b[i,j] \getsr \Ber(1/2)$ for $i,j \in \{1,\cdots,l\}$ \Comment perform $l^2$ unbiased independent coin flips
  \For{$i \gets 1$ to $l$}
    \If{$b[i,1]=b[i,2]=\cdots=b[i,k]=1$}
      \State $\chi \gets \chi \cup \{a_i\}$
    \Else
      \State $\chi \gets \chi - \{a_i\}$
    \EndIf
    \If{$|\chi| = n$}
      \State $\chi \gets \{a \in \chi \;|\; b[\textrm{last{\textunderscore}index}(a),k+1] = 1\}$
      \State $k \gets k+1$
    \EndIf
    \If{$|\chi| = n$}
      \State \Return $\bot$
    \EndIf
  \EndFor
  \State \Return $2^k |\chi|$ \Comment estimate cardinality of $A$
  \end{algorithmic}
\end{algorithm}
To roughly explain, how tail bounds can be derived for this new variant:
It is possible to union bound the probability that the estimate exceeds the desired interval, with the probability of the event in conjunction with a specific value of $k$, which can be bounded by the probability that the number of stream elements with whose associated coin flips start with $k$ ones, is outside of $2^{-k} |A| (1 \pm \varepsilon)$.
This is explained in more detail by Chakraborty et al.~\cite{chakraborty2022}.
